A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-07-12 evaluation.tex
more

To evaluate the effectiveness of \Calculus{} in an actual string solver,
\Catra{} was experimentally integrated as the Parikh automata product solver
into the string solver \Ostrich{}~version~1.3. \Ostrich{} is an independently
developed solver that participated in the recent SMT-COMP~2023\footnote{\url{https://smt-comp.github.io/2023/participants/ostrich}},
 winning the single-query track
for quantifier-free strings (\texttt{QF\_S}), as well as dominating
other solvers on unsatisfiable string benchmarks~\cite{smt-comp-23}.

At SMT-COMP, \Ostrich{} competed using three different back-ends:
\begin{itemize}
\item \textbf{BW-Str:} a backward-propagation-based solver, not utilising
  Parikh automata~\cite{ostrich}.
\item \textbf{ADT-Str:} a solver based on algebraic data-types.
\item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
  algorithm~\cite{ostrichplus}., using Parikh automata and a variant
  of our baseline method~\cite{generate-parikh-image}.
\end{itemize}

@@

As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
which we picked 1000~problems uniformly at random. We ran the
different back-ends on those 1000~problems with a 120s timeout.
%
%The \texttt{QF\_S} track contains no Parikh automata intersection problems,
%however. For that, we need to look to the linear integer arithmetic
%(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
%\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
%\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
%competition. In this section, we evaluated a version of \Ostrich{} using
%\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
%on \numprint{1000}~randomly selected instances from the PyEx benchmarks. 
The results are shown in \cref{tab:solve-status-smt-comp}.
%
% comparison \Fudge{to the other provers of the 2023
%competition on the same benchmarks, as well as \Ostrich{} using the baseline
%back-end described in \cref{sec:implementing-baseline}}.

We include this section to show that \Calculus{} can be used to enhance the
performance of an automaton-based string solver with cost-enriched automata like
\Ostrich{}. The results should be considered as preliminary, however, 
as we believe that a deeper integration of our \Catra{} solver into
string solvers can lead to significant performance gains.

>>>>

To evaluate the effectiveness of \Calculus{} in an actual string solver,
\Catra{} was experimentally integrated as the Parikh automata product solver
into the string solver \Ostrich{}~version~1.3. \Ostrich{} is an independently
developed solver that participated in the recent SMT-COMP~2023\footnote{\url{https://smt-comp.github.io/2023/participants/ostrich}},
 winning the single-query track
for quantifier-free strings (\texttt{QF\_S}), as well as dominating
other solvers on the complete set of
 unsatisfiable string benchmarks~\cite{smt-comp-23}.

 At SMT-COMP, \Ostrich{} competed using different
 (automata-based) back-ends:
\begin{itemize}
\item \textbf{BW-Str:} a backward-propagation-based solver, not utilising
  Parikh automata~\cite{ostrich}.
%\item \textbf{ADT-Str:} a solver based on algebraic data-types.
\item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
  algorithm~\cite{ostrich-plus}, using Parikh automata and the
  encoding from~\cite{generate-parikh-image}.
\end{itemize}

@@

As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
which we picked 1000~problems uniformly at random. We ran
different combinations of the
back-ends on those 1000~problems with a 120s timeout.
%
%The \texttt{QF\_S} track contains no Parikh automata intersection problems,
%however. For that, we need to look to the linear integer arithmetic
%(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
%\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
%\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
%competition. In this section, we evaluated a version of \Ostrich{} using
%\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
%on \numprint{1000}~randomly selected instances from the PyEx benchmarks. 
The results are shown in \cref{tab:solve-status-smt-comp}.
%
% comparison \Fudge{to the other provers of the 2023
%competition on the same benchmarks, as well as \Ostrich{} using the baseline
%back-end described in \cref{sec:implementing-baseline}}.

As can be seen from the table, integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. The
increase is more relevant for the unsatisfiable problems, given that
\Ostrich{} is a state-of-the-art solver in particular on such benchmarks.

The results show that \Calculus{} can be used to enhance the
performance of an automaton-based string solver.
The results should be considered as preliminary, however, 
as we believe that a deeper integration of our \Catra{} solver into
string solvers can lead to significant performance gains.
